---
title: "Housing Prices - Ames Iowa"
author: "Vy Nguyen"
date: "January 23, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Libraries

```{r}
rm(list=ls())
library(dplyr)
library(tidyr)
library(glmnet)
library(caret)
library(mice) # Impute missing values
library(VIM) # Visualizing missing data

```

Overview understanding of the data

```{r }
train <- read.csv("train.csv")
test <- read.csv("test.csv")
names(train)
head(train)
summary(train)
str(train)
```

## Visualizations

### Visualization for missing data

```{r}
pMiss <- function(x){sum(is.na(x))/length(x)*100} # Calculate the percentage of missing values
missVar <- apply(train,2,pMiss)
missVar
aggr_plot <- aggr(train, col=c('navyblue','red'), numbers=TRUE, sortVars=TRUE, labels=names(train), cex.axis=.7, gap=3, ylab=c("Histogram of missing data","Pattern"))

```

The data contains few missing values as we note some variables as follows:
* `BsmtQual`,`BsmtCond`, `BsmtExposure`, `BsmtFinType1`, `BsmtFinType2`
* `GarageType`, `GarageYrBlt`, `GarageFinish`, `GarageQual`, `GarageCond`
* `Alley`, `FireplaceQu`, `PoolQC`, `Fence`,`MiscFeature`

whose `NA` values simply mean the house does not have that feature/facility.

Variables with true missing values are below.
* `LotFrontage`: Linear feet of street connected to property (18% missing values)
* `MasVnrType`, `MasVnrArea`: Masonry Veneer type houses (0.5% missing values) - consider removing obs
* `Electrical`: Electrical system (0.07% missing values) - consider removing obs

`LotFrontage` has the most missing values. The other two variables have very few missing values. We could consider using imputation methods to impute the missing values.

### Visualizations for numeric variables

```{r}


```


### Visualizations for categorical variables

```{r}


```

### Building the models

#### The Lasso

```{r}
## Removing and/or Imputing missing values
train <- subset(train,!(is.na(MasVnrType)|is.na(Electrical))) # Removing observations with NAs values
noneCol <- c(7, 31:34, 36,58:59,61,64:65,73:75) # Categorical variables with NA for 'None' value
zeroCol <- c(60) # Continuous variables with NA to be replaced with 'Zero' value

length(noneCol) + length(zeroCol)
sum(apply(train[,noneCol],2,is.factor)) == 0 # Check if they are factor variables

for (i in noneCol){
  train[,i] <- addNA(train[,i]) #Add NA as a factor variable
}

# train[,noneCol] <- apply(train[,noneCol],2,addNA) 
train[,noneCol] <- factor(train[,noneCol])
#x <- train[,zeroCol]
#x[is.na(x)] <- 0
#train[,zeroCol] <- x
train[is.na(train[,zeroCol]),zeroCol] <- 0 #Assigning zero values to zeroCol

tmp <- mice(train,method="cart")

for (i in (1:5)){
  completed.i <- complete(tmp,i)
  write.table(completed.i, "impute.txt",
            sep="\t", row.names=FALSE)
  imp.i = read.table("impute.txt",sep="\t",header = TRUE)
}

```

